%%%%%%%%%%%%%%%%%%%%%%%%%%% mainthesis.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%
% Template for University of Maryland Thesis and Dissertation files  %
% Written by   Michael Jarret                                        %
%              Perimeter Institute of Theoretical Physics            %
%              Waterloo, Ontario, Canada                             %
%              June 25, 2017                                         %
% First version written Feb, 2016 while a student at                 %
%       Joint Center for Quantum Information and Computer Science    %
%       University of Maryland, College Park                         %
%       Feb, 2016                                                    %
% Modified: June 25, 2017 by Michael Jarret                          %
% Based upon: UMD Dissertation Template by Dorothea F Brosius        %
%             Institute for Electronics and Applied Physics          %
%             University of Maryland, College Park                   %
%             Version from June 2015                                 %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%t

\documentclass[letterpaper,12pt]{umd-thesis}  
\usepackage{mathptmx} % This is a close match to Timesop New Roman, which is recommended by the grad school

% hyperref automatically hyperlinks references
\usepackage{hyperref,xcolor}

% graphics packages
\usepackage{graphicx,epstopdf}			

% path to graphics files
\graphicspath{{Figures/}} 

% Some math packages
\usepackage{amsmath,amsthm,amsfonts,amssymb,mathrsfs,dsfont,breqn} 

% Cleveref MUST come after ams packages if AMS packages are used.
\usepackage{cleveref}  

% Natbib citation package is probably most elegant
\usepackage[numbers,sort&compress]{natbib}
\bibliographystyle{unsrtnat}

%\usepackage[sorting=none]{biblatex}
%\bibliography{main.bib}


\usepackage{afterpage}
\usepackage{times}
\usepackage{enumitem}
\usepackage{subcaption}
\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{mathtools}
\usepackage{pdflscape}
\usepackage{tabularx}
\usepackage{adjustbox}


\newenvironment{changemargin}[2]{%
  \begin{list}{}{%
    \setlength{\topsep}{0pt}%
    \setlength{\leftmargin}{#1}%
    \setlength{\rightmargin}{#2}%
    \setlength{\listparindent}{\parindent}%
    \setlength{\itemindent}{\parindent}%
    \setlength{\parsep}{\parskip}%
  }%
  \item[]}{\end{list}}

% These are my own custom latex commands, uncomment if unwanted
\input{custom_commands.tex}

% Custom formatting options
\hypersetup{linktocpage, linkbordercolor=blue}  % Blue outlined hyperlinks in TOC

% This changes the numbering style to (chapter).(equation/thm/etc.) 
% Note that this is an amsmath command and depends upon the inclusion
% of the amsmath pacakage
\numberwithin{equation}{chapter}  
\numberwithin{thm}{chapter}
\numberwithin{lem}{chapter}
\numberwithin{cor}{chapter}
\numberwithin{definition}{chapter}

% Plain is the standard choice
%\bibliographystyle{plain}

% Everything here is reasonably obvious
\title{Purity  Monitoring  Techniques  and  Electronic Energy Deposition Properties  in Liquid Xenon Time Projection  Chambers}
\author{Jon Balajthy}
\department{Department of Physics}
\gradyear{2018}
\degree{Doctor of Philosophy}

% Your {advisor}[his department][his institution]
\principaladvisor{Professor Carter Hall}[University of Maryland] 

% Comment out if no co-advisor
%\coadvisor{Professor Alexey Gorshkov}[Department of Physics]
% Uncomment the following line if your co-advisor is your chair

% Uncomment the following line if someone other than your principal advisor or co-advisor is your chair and add his/her details
%\committeechair{Professor Carter Hal}[Department of Physics]

% Add any additional committee members in alphabetical order below
\firstreader{Professor Elizabeth Beise}

\secondreader{Professor Alberto Belloni}

\thirdreader{Professor Xiangdong Ji}

\fourthreader{Professor Coleman Miller}

% The options below should be self explanatory, uncomment as needed
% \coadvisorischair % Only use this if you did not add a committee chair
% \copyrightfalse
% \tablespagefalse
%\figurespagefalse
% \contentspagefalse
% \bibpagefalse

\begin{document}

% Type the abstract below
\begin{abstract}
Currently, one of the most well motivated models of dark matter is the weakly interacting massive particle (WIMP), and the detector technology that is in the best position to observe these WIMPs is the liquid xenon time projection chamber (TPC). As liquid-xenon WIMP detectors grow larger and more sensitive, the requirements placed on their signals and backgrounds become more and more stringent. In this document, we develop a technique for measuring the concentration of one of the more problematic radioactive backgrounds, $^{85}$Kr, in xenon. We show that with this technique we are able to detect natural krypton concentrations down to 7.7 $\pm$2.0 parts per quadrillion (ppQ). On the signals side, we provide a measurement of the charge and light yields of beta deposits in liquid xenon. For these measurements, we use $^{14}$C calibration data collected in the LUX detector after the 2014-2016 WIMP-search run was completed. We also look for a non-statistical shape factor in the $^{14}$C spectrum. We show that the spectral shape observed in the LUX calibration data is consistent with a purely statistical shape, and disagrees with a recent measurement by Kuzminov et al by 1.8-$\sigma$. However, pathologies in the LUX signals prevent us from making any strong claims on this topic.

\end{abstract} 

% This is the frontmatter and can include the preface, foreword, dedication, and acknowledgements. Any
% of these can be commented out if unwanted.
\begin{preliminary}

    %\begin{preface}
       % \lipsum[1]
    %\end{preface}

    %\begin{foreword}
        %\lipsum[1]
    %\end{foreword}


	%\begin{dedication}
		
	%\end{dedication}

	\begin{acknowledgements}	
	I would like to take this opportunity to thank everyone who has helped me in completing the work for this dissertation. I began working on LUX in 2010 as an undergraduate for my current advisor, Carter Hall. His help and guidance have pushed me to be a better experimenter and have given me the tools to do so. I want to thank him for all of the support he has given me over the past 8 years. 
	
	My colleagues in the Experimental Nuclear Physics group here at UMD have all helped my to develop my knowledge and skills along the way. In particular, I would like to thank Attila and Richard for helping to develop both my analysis skills and my knowledge of building and operating ultra-high purity plumbing systems. Tim and John, I want to thank you for helping me to hone my skills and for helping me when a particularly challenging problem would arise. I also need to thank Tom, Simon, Clayton, and Ruey for sharing some of their knowledge and expertise with me along the way.
	
	I owe a great deal of gratitude to my collaborators at SLAC. Dan, Tom, Christina, Wing, TJ, Eli, Wei, Tomasz, Shaun, Gabe, and Kelly, thank you for all of the support you have given me. Thank you especially for bring a couch in for me to lay on during the long nights spent developing the sampling system. 
	
	I also need to thank all of my fellow LUXers, all of whom devoted part of themselves to building and operating this amazing experiment. In particular I want to thank Elizabeth, Mongkol, Markus, Lucie, Dev, Evan, Curt, Simon, Mark, Aaron, and everyone else who made working on site in Lead such an amazing experience. 
	
	To all of the personnel at SURF, thank you for keeping our experiment running even when we needed immediate access to the Davis campus at 1am on a Saturday night. To Robyn, thank you for making the 4850 so sunny. 
	
	I would also like to thank Matthew for helping me to understand the recombination model and for helping me to compile the all the puzzle pieces I needed. You helped make clear an otherwise opaque topic.
	
	To my family, who have supported and provided much needed-distractions for me along the way, thank you. My sister, my cousins, my Aunts and Uncles, and my Mother, Sister, and soon-to-be brother in law have all helped me to stay sane over the years. I especially want to thank my parents and grandparents, who made me the person I am today and have never stopped giving me their love and support. I love you all.
	
	I finally want to thank my beautiful wife Caroline. You have constantly driven me to be better and to do more. Thank you for withstanding the late nights, the months apart, and the constant talk about physics. Without your love and support, I never would have made it through. 
	
	To everyone else who has helped me over the years: thank you, and I'm sorry I failed to mention you here.
	\end{acknowledgements}
	
%% if you need a "List of Symbols or Abbreviations" look into
% gatech-thesis-gloss.sty.
\end{preliminary}


\include{chapters/Advancement_paper}
\include{chapters/LXe_TPCs}
\include{chapters/Sampling_Manual}
\include{chapters/Post_Run04_Calibrations}
\include{chapters/C14_Calibration}



\chapter{Conclusions}
Direct detection dark matter experiments are fast approaching the solar neutrino limit, with LZ projected to reach within about an order of magnitude of this floor. To reach this limit, experimenters must understand the backgrounds of their detectors better than ever before. In chapter \ref{chap:sampling} of this document, we have presented a technique for monitoring the amount of the radioactive $^{85}$Kr and $^{39}$Ar isotopes dissolved in a xenon detector. We showed that this technique is able to measure the concentration of natural krypton down to 7.7 $\pm$2.0 ppQ (grams of krypton per gram of xenon). This corresponds to a $^{85}$Kr activity of about 0.002 events years$^{-1}$ kg$^{-1}$ keV$^{-1}$, which is below the LZ goal. For in-situ measurements which will benefit from an impurity amplification due to the solubility of krypton in liquid xenon, we find that the method laid out in chapter \ref{chap:sampling} is expected to achieve the sensitivity requirement of a post-LZ era dark matter experiment.

In order to obtain a trustworthy measurement of the $^{14}$C shape factor, a precise and physically motivated model of the detector resolution is required. This model must include any time, event rate, or position dependence of the detector response. Additionally, an equivalent model of the event selection acceptance must also be known. In section \ref{sec:shape_fit}, we attempted to make such a measurement and found that we were not able to reproduce the results from \cite{C14_Kuzminov}. We instead found that our measured spectrum is consistent with a purely statistical shape. Pathologies in the $S2$ signal prevent us from making any strong claims about this topic, however. The pathologies, which we described and modeled in section \ref{sec:s2tails}, may in fact limit the ability of any 2-phase xenon TPC from making meaningful nuclear physics measurements that require stable, keV-level energy resolution over a wide range of energies and over extended periods of time. For such a measurement, it may be that the single phase detectors possess inherent advantages over their two-phase counterparts. Single phase detectors measure the charge signal in the liquid phase and therefore do not have the added detector fluctuations that are introduced due to electron extraction efficiency. 

These next generations of dark matter detectors will also need a field and energy dependent understanding of the energy deposition properties in liquid xenon. For LZ, knowledge of the beta yields will be vitally important for calculations of the discrimination between NR and ER events, because its ER backgrounds are dominated by beta decays from radon daughters\cite{lz_tdr}. In section \ref{sec:NESTbeta} we presented an empirical model of the recombination process for beta deposits in liquid xenon. To fit this model, we take into account data from 8 different experiments, at 18 different field values from 0 to 4050 V/cm, and spanning energies from 0.27 keV up to 1 MeV. This model can be further developed by using more physically motivated fitting functions such as the Thomas-Imel model presented in section \ref{sec:libnest}. As it stands, it is a powerful tool that can be used to predict the response of a xenon detector to an ER beta event.





\appendix
\include{chapters/Appendix_results}

\bibliographystyle{unsrtnat}
\bibliography{main.bib}


%\printbibliography


\end{document}